\documentclass[11pt]{article}

% ---------------------------------------------------------------------------
% FAITH-SAE Roadmap
% Author: Rajia Rani
% No lab/team affiliation. For research and educational purposes only.
% ---------------------------------------------------------------------------

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=0.85in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage[hidelinks]{hyperref}
\hypersetup{colorlinks=true, linkcolor=accent!70!black, urlcolor=accent!70!black,
            citecolor=accent!70!black}

\tcbuselibrary{skins,breakable}

% ----- palette -------------------------------------------------------------
\definecolor{accent}{RGB}{34,94,168}      % deep blue
\definecolor{accent2}{RGB}{0,128,128}     % teal
\definecolor{snapbg}{RGB}{232,240,250}
\definecolor{snaprule}{RGB}{34,94,168}
\definecolor{insightbg}{RGB}{235,246,242}
\definecolor{insightrule}{RGB}{0,128,128}
\definecolor{delivbg}{RGB}{252,244,228}
\definecolor{delivrule}{RGB}{198,138,28}
\definecolor{scopebg}{RGB}{244,238,248}
\definecolor{scoperule}{RGB}{120,80,160}
\definecolor{checkbg}{RGB}{238,243,236}
\definecolor{checkrule}{RGB}{80,140,70}

% ----- header / footer -----------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\lhead{\small\textsc{FAITH-SAE Roadmap}}
\rhead{\small Rajia Rani}
\lfoot{\small 2026-06-20 \;$\cdot$\; For research and educational purposes only.}
\rfoot{\small\thepage}

% ----- section styling -----------------------------------------------------
\titleformat{\section}{\Large\bfseries\color{accent}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\large\bfseries\color{accent!85!black}}{\thesubsection}{0.5em}{}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.35em}
\titlespacing*{\section}{0pt}{1.0em}{0.4em}
\titlespacing*{\subsection}{0pt}{0.7em}{0.3em}

% ----- reusable boxes ------------------------------------------------------
\newtcolorbox{snapbox}[1][]{colback=snapbg,colframe=snaprule,
  boxrule=0.8pt,arc=2pt,left=8pt,right=8pt,top=6pt,bottom=6pt,breakable,#1}
\newtcolorbox{insightbox}[1][]{colback=insightbg,colframe=insightrule,
  boxrule=0.8pt,arc=2pt,left=8pt,right=8pt,top=6pt,bottom=6pt,breakable,
  title=#1,fonttitle=\bfseries\color{white},colbacktitle=insightrule}
\newtcolorbox{delivbox}[1][]{colback=delivbg,colframe=delivrule,
  boxrule=0.8pt,arc=2pt,left=8pt,right=8pt,top=4pt,bottom=4pt,breakable,
  title=#1,fonttitle=\bfseries\color{white},colbacktitle=delivrule}
\newtcolorbox{scopebox}[1][]{colback=scopebg,colframe=scoperule,
  boxrule=0.8pt,arc=2pt,left=8pt,right=8pt,top=4pt,bottom=4pt,breakable,
  title=#1,fonttitle=\bfseries\color{white},colbacktitle=scoperule}
\newtcolorbox{checkbox}[1][]{colback=checkbg,colframe=checkrule,
  boxrule=0.8pt,arc=2pt,left=8pt,right=8pt,top=4pt,bottom=4pt,breakable,
  title=#1,fonttitle=\bfseries\color{white},colbacktitle=checkrule}

\newcommand{\cbox}{$\square$\;}

% ===========================================================================
\begin{document}

\begin{center}
  {\LARGE\bfseries\color{accent} FAITH-SAE: Are Sparse-Autoencoder Concept\\[2pt]
   Directions in Vision Models Causally Faithful\\[2pt]
   Under Distribution Shift?}\\[8pt]
  {\large 8-Week Research Roadmap}\\[5pt]
  {\normalsize Rajia Rani \;$\cdot$\; \texttt{}}\\[3pt]
  {\small 2026-06-20 \;$\cdot$\; For research and educational purposes only.}
\end{center}

\vspace{0.3em}
\hrule
\vspace{0.6em}

% ===========================================================================
\section{What This Project Is About}

\begin{snapbox}
\textbf{\color{accent} Project Snapshot}\\[4pt]
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}>{\bfseries}p{0.20\linewidth} p{0.74\linewidth}@{}}
Building blocks &
  Steering methods over a \emph{frozen} CLIP ViT-B/16 + TopK sparse autoencoder:
  \texttt{naive\_steer} (off-manifold activation-addition), \texttt{random\_steer},
  \texttt{clamp\_steer}, and \texttt{onmanifold\_steer} (project the edit onto the
  top-$r$ real-image subspace, \emph{ours}).\\
Baselines &
  Supervised TCAV-style concept-direction steering (quality reference), naive
  off-manifold activation-addition (main competitor), random-direction (null/sanity),
  raw-clamp --- all at \textbf{matched steering strength}.\\
Benchmarks &
  Clean ImageNet-val $\rightarrow$ ImageNet-R $\rightarrow$ ImageNet-Sketch
  $\rightarrow$ ImageNet-C (severity 1--5) $\rightarrow$ ObjectNet.\\
Metrics &
  Monotonicity, specificity, sufficiency $\rightarrow$ \textbf{Causal Faithfulness
  Score (CFS)} $\in[0,1]$; off-manifold residual; $\Delta$CFS-per-shift degradation
  slope with bootstrap CIs over concepts.\\
Big idea &
  A steering \emph{method} with no metric is unfalsifiable; a \emph{metric} applied to
  naive edits just measures artifacts. Combine them: measure CFS of on-manifold
  steering \emph{across the OOD ladder}. Each mechanism cures the other's blind spot.\\
Target &
  On-manifold steering achieves higher CFS than naive / random / clamp at matched
  strength on clean images, approaches the supervised reference, and degrades more
  gracefully under shift --- locating the collapse knee.\\
\end{tabular}
\end{snapbox}

Sparse autoencoders (SAEs) decompose a frozen backbone's dense, polysemantic
activations into a large dictionary of sparse, human-nameable \emph{concept
directions}. In vision this has produced striking demonstrations --- directions that
appear to encode ``stripes,'' ``dog face,'' or ``snowy scene.'' But a picture that
\emph{looks} interpretable is not the same as a direction that is \emph{causally real}:
if we push the ``dog face'' direction up, does behaviour move smoothly and specifically
the way the concept's name promises, or have we merely produced a plausible artifact?

This project studies two ingredients usually examined apart.
\textbf{Mechanism A --- on-manifold steering (the method)} forces a chosen SAE concept
up or down inside the frozen backbone, but \emph{constrains} the edit: instead of the
raw activation-addition $\Delta$, it adds the projection $P_M\!\cdot\!\Delta$ onto the
top-$r$ principal directions of \emph{real}-image activations, so the steered
activation stays in the region the model was trained on. The edit is realistic,
decodable, and non-destructive. Its weakness alone: a steering method with no metric is
unfalsifiable. \textbf{Mechanism B --- the CFS metric plus OOD stress (the measuring
stick)} is a single score in $[0,1]$ asking whether an edit is \emph{monotone} (knob up
$\rightarrow$ readout up smoothly), \emph{specific} (only the target concept moves), and
\emph{sufficient} (effect size matches the claimed magnitude) --- re-measured as images
get harder. Its weakness alone: on naive steering it just quantifies off-manifold
artifacts.

\begin{insightbox}[The opening]
The unoccupied space is the \textbf{combination}: measure the CFS of on-manifold
steering \emph{across} the OOD sweep (clean $\rightarrow$ R $\rightarrow$ Sketch
$\rightarrow$ C $\rightarrow$ ObjectNet). On-manifold steering gives the metric
something \emph{real} to measure; the metric proves the on-manifold edit is faithful
and tracks exactly where it breaks. The CFS-versus-shift curve is the paper's answer.
Prior work shows SAEs \emph{find} vision concepts and that you \emph{can} steer them,
and separately warns that off-manifold steering is unreliable --- but no one has a
\textbf{quantitative faithfulness benchmark for vision SAE features}, and no one has
tested faithfulness \textbf{under distribution shift}.
\end{insightbox}

\begin{scopebox}[Scope \& Realism Note]
This is a \emph{controlled, reproducible} interpretability study, not a leaderboard
chase. We use one small, public, frozen backbone (CLIP ViT-B/16) and a fixed public
data budget (ImageNet-val and its shifted variants), so every comparison holds the
backbone, SAE, and steering strength constant and changes only the steering
\emph{method} or one design knob. The headline quantity --- the \textbf{Causal
Faithfulness Score}, a backbone- and library-independent number per concept and per
shift level (the analog of ``bytes/token'') --- is reported alongside wall-clock so
results transfer beyond our setup. Every figure ships with \emph{illustrative}
placeholder numbers until the real CLIP + ImageNet-shift runs land at Milestone~3.
\end{scopebox}

% ===========================================================================
\section{Research Questions \& Timeline}

\textbf{RQ1 (headline --- on-manifold vs.\ the field).}
Does on-manifold projected steering achieve a higher Causal Faithfulness Score (CFS)
than naive off-manifold activation-addition steering --- and than random-direction,
raw-clamp, and supervised concept-direction (TCAV-style) steering --- at
\textbf{matched steering strength} on clean, in-distribution images?

\textbf{RQ2 (core knob --- decomposition \& design).}
How does CFS \emph{decompose} into monotonicity / specificity / sufficiency, what
fraction of discovered concepts steer reliably (the field's ``only $\sim$10--15\%''
claim), and how does the faithfulness optimum depend on the two design knobs ---
\textbf{steering strength} $s$ and \textbf{manifold-projection rank} $r$?

\textbf{RQ3 (the sweep --- collapse knee).}
Does CFS \emph{survive distribution shift} across clean ImageNet $\rightarrow$
ImageNet-R $\rightarrow$ ImageNet-Sketch $\rightarrow$ ImageNet-C (corruption-severity
dial) $\rightarrow$ ObjectNet, and \textbf{where is the collapse knee} --- i.e.\ what
is the $\Delta$CFS-per-shift-level degradation slope, and does on-manifold steering
degrade more gracefully than naive?

\vspace{0.4em}
\renewcommand{\arraystretch}{1.3}
\begin{center}
\begin{tabular}{@{}p{0.20\linewidth} p{0.12\linewidth} p{0.60\linewidth}@{}}
\toprule
\textbf{Milestone} & \textbf{Weeks} & \textbf{Key Deliverable(s)}\\
\midrule
M1 --- Literature Review \& Block Design & W1--2 &
  Annotated 12-paper review; CFS definition fixed; the four steering variants and the
  evaluation protocol designed; Deliverable~\#1.\\
M2 --- Data Pipeline \& Implementation & W3--4 &
  CLIP ViT-B/16 + TopK SAE pipeline; the OOD data ladder loaded; all variants
  implemented and validated on the synthetic planted-concept task; Deliverable~\#2.\\
M3 --- Training \& Ablation Experiments & W5--6 &
  Full CFS measurements across variants and the OOD ladder; ablations A1--A5; error
  taxonomy populated; \texttt{results/metrics\_all.csv}; Deliverable~\#3.\\
M4 --- Analysis, Insights \& Manuscript & W7--8 &
  9 figures; insight report answering RQ1--RQ3; practitioner decision table;
  IEEE-conference manuscript draft; Deliverable~\#4.\\
\bottomrule
\end{tabular}
\end{center}

% ===========================================================================
\section{Milestone 1: Literature Review \& Block Design}

\textbf{Objectives (three pillars).}
\begin{enumerate}[leftmargin=1.4em,itemsep=2pt]
  \item \textbf{Concepts.} Master how vision SAEs discover concept directions (TopK vs.\
    L1 dictionaries, monosemanticity, the polysemantic tail) so we know what we are
    steering and why most directions are not clean knobs.
  \item \textbf{Steering \& faithfulness.} Pin down activation-addition / representation
    engineering, the on-manifold constraint $P_M\!\cdot\!\Delta$, and what it means for
    an edit to be causally faithful (monotone, specific, sufficient) --- then formalize
    CFS as a single conjunctive score.
  \item \textbf{Robustness.} Understand the distribution-shift ladder (renditions,
    sketch, graded corruption, real-world ObjectNet shift) so the OOD sweep is a
    principled stress test, not an arbitrary sample.
\end{enumerate}

\renewcommand{\arraystretch}{1.2}
\begin{center}\small
\begin{tabular}{@{}p{0.24\linewidth} p{0.40\linewidth} p{0.30\linewidth}@{}}
\toprule
\textbf{Component} & \textbf{Mechanism} & \textbf{Role}\\
\midrule
TopK Sparse Autoencoder & Sparse dictionary over frozen ViT activations; keep top-$k$
  features active per token & Discovers the concept directions we steer\\
Concept direction $d$ & A single SAE decoder column (a feature's direction in
  activation space) & The knob we push up/down\\
On-manifold projection $P_M$ & $P_M=U_rU_r^{\!\top}$ onto top-$r$ real-image PCA
  subspace; $a'=a+s\,(P_M\Delta)$ & Keeps the edit realistic / decodable (the method)\\
Steering strength $s$ & Scalar multiplier on the (projected) edit & The intensity knob;
  too large $\Rightarrow$ saturation\\
CFS $=\mathrm{HM}$(mono, spec, suff) & Harmonic mean $\Rightarrow$ conjunctive: faithful
  only if \emph{all three} hold & The single contributed measuring stick\\
OOD ladder & clean $\to$ R $\to$ Sketch $\to$ C(1--5) $\to$ ObjectNet & The shift axis;
  yields the CFS-vs-severity curve\\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Reading list (12 papers)}
\begin{enumerate}[leftmargin=1.6em,itemsep=0pt]
  \item Cunningham et al.\ 2023 --- \emph{Sparse Autoencoders Find Highly Interpretable
    Features} --- arXiv:2309.08600.
  \item Bricken et al.\ 2023 --- \emph{Towards Monosemanticity: Decomposing Language
    Models with Dictionary Learning} (Anthropic).
  \item Gao et al.\ 2024 --- \emph{Scaling and Evaluating Sparse Autoencoders} (TopK
    SAE) --- arXiv:2406.04093.
  \item Templeton et al.\ 2024 --- \emph{Scaling Monosemanticity} (Anthropic).
  \item Turner et al.\ 2023 --- \emph{Activation Addition / ActAdd: Steering Language
    Models} --- arXiv:2308.10248.
  \item Zou et al.\ 2023 --- \emph{Representation Engineering: A Top-Down Approach to AI
    Transparency} --- arXiv:2310.01405.
  \item Kim et al.\ 2017 --- \emph{Interpretability Beyond Feature Attribution (TCAV)}
    --- arXiv:1711.11279.
  \item Radford et al.\ 2021 --- \emph{Learning Transferable Visual Models from Natural
    Language (CLIP)} --- arXiv:2103.00020.
  \item Dosovitskiy et al.\ 2020 --- \emph{An Image is Worth 16x16 Words (ViT)} ---
    arXiv:2010.11929.
  \item Hendrycks et al.\ 2020 --- \emph{The Many Faces of Robustness (ImageNet-R)} ---
    arXiv:2006.16241.
  \item Hendrycks \& Dietterich 2019 --- \emph{Benchmarking Neural Network Robustness
    (ImageNet-C)} --- arXiv:1903.12261.
  \item Wang et al.\ 2019 --- \emph{Learning Robust Global Representations
    (ImageNet-Sketch)} --- arXiv:1905.13549.
\end{enumerate}
The manuscript expands this to $\sim$24 entries with ObjectNet (Barbu et al.\ 2019),
Causal Scrubbing (Chan et al.\ 2022), Makelov et al.\ (2405.08366), Rajamanoharan
et al.\ \emph{JumpReLU / Gated SAEs} (2407.14435 / 2404.16014), Marks et al.\ \emph{Sparse
Feature Circuits} (2403.19647), Fel et al.\ (vision dictionary learning), Hooker et al.\
\emph{ROAR} (1806.10758), and Geirhos et al.\ \emph{Texture-vs-Shape Bias} (1811.12231).

\begin{insightbox}[Core Research Insight]
Faithfulness is \textbf{conjunctive}, not additive. A steered concept is trustworthy
only if it is \emph{monotone and specific and sufficient at once} --- a near-zero on
any single axis (e.g.\ an edit that moves the readout but also drags every off-target
probe) should collapse the whole score. We therefore define CFS as the \emph{harmonic
mean} of the three components, so one weak axis dominates. The on-manifold projection
$P_M\Delta$ is what gives that conjunction a chance to hold: by keeping the edit inside
the real-image subspace, it preserves specificity and decodability that naive
off-manifold addition destroys --- and we expect that advantage to widen, not shrink,
as inputs move out of distribution.
\end{insightbox}

\subsection*{Evaluation plan}
\textbf{Metrics.} Per concept: monotonicity (Spearman correlation between the steering
knob and a held-out concept readout), specificity ($1-$ normalized drift of off-target
linear probes), sufficiency (standardized Cohen's-$d$ effect size at full knob vs.\ the
concept's claimed magnitude); composite CFS $=\mathrm{HM}(\cdot)$; off-manifold
residual $\lVert\Delta-P_M\Delta\rVert/\lVert\Delta\rVert$; and the OOD degradation
slope $\Delta\mathrm{CFS}/\Delta\text{shift}$ with bootstrap CIs over concepts.
\textbf{Protocols.} Matched steering strength across all methods; held-out readout
probes never seen during direction selection; the same selected concept set evaluated
at every shift level. \textbf{Baselines.} Supervised TCAV-style (quality reference),
naive off-manifold (main competitor), random-direction (null), raw-clamp.

\begin{delivbox}[Deliverable \#1 --- Review \& Design]
A written literature review of the 12 papers; a frozen, equation-level definition of
CFS and the on-manifold edit $a'=a+s(P_M\Delta)$; the four steering variants
(\texttt{naive\_steer}, \texttt{random\_steer}, \texttt{clamp\_steer},
\texttt{onmanifold\_steer}) and the supervised reference specified; the evaluation
protocol (metrics, matched-strength rule, OOD ladder) fixed.
\end{delivbox}
\begin{checkbox}[Acceptance Check]
CFS and $P_M\Delta$ are written down unambiguously; each of the 5 baselines/variants has
a one-line precise definition; the protocol guarantees every method is compared at
matched strength on the \emph{same} concepts; a reviewer can restate the open gap
(no quantitative vision-SAE faithfulness benchmark; none tested under shift) in one
sentence.
\end{checkbox}

% ===========================================================================
\section{Milestone 2: Data Pipeline \& Implementation}

\renewcommand{\arraystretch}{1.2}
\begin{center}\small
\begin{tabular}{@{}p{0.20\linewidth} p{0.13\linewidth} p{0.34\linewidth} p{0.26\linewidth}@{}}
\toprule
\textbf{Dataset} & \textbf{Size} & \textbf{Description} & \textbf{Use}\\
\midrule
ImageNet-val & 50k & Clean in-distribution validation images & Quality reference: clean
  CFS, concept readouts, SAE training source\\
ImageNet-R & 30k & Renditions / abstractions (art, cartoons, sculpture) & Does
  faithfulness survive style/domain shift\\
ImageNet-Sketch & 50k & Black-and-white sketches; texture removed & Concept readout
  without texture cues (hard stress)\\
ImageNet-C & 50k$\times$ & 15 corruptions $\times$ severity 1--5 (the dial) & Graded
  OOD: the headline CFS-vs-severity curve\\
ObjectNet & 50k & Real-world pose / background / viewpoint controls & Hardest shift;
  the faithfulness floor\\
\bottomrule
\end{tabular}
\end{center}

\textbf{Tasks.}
\begin{enumerate}[leftmargin=1.5em,itemsep=1pt]
  \item Load frozen CLIP ViT-B/16; extract \& cache patch (and CLS) activations at the
    chosen layer for ImageNet-val.
  \item Train a TopK SAE; build the concept dictionary; estimate the top-$r$ real-image
    PCA subspace $U_r$ for $P_M$.
  \item Implement the four steering variants behind one registry
    (\texttt{src/blocks/\_\_init\_\_.py}), plus the supervised TCAV-style reference and
    the held-out readout / off-target probes.
  \item Implement OOD loaders for R / Sketch / C(1--5) / ObjectNet with a fixed, matched
    evaluation protocol across shifts.
  \item Wire \texttt{cfs\_score}, \texttt{onmanifold\_projection\_residual}, and the
    \texttt{faithfulness(variant,cfg)} dispatcher so \texttt{run\_experiments} and the EDA
    notebook share one scoring model; emit \texttt{cfs} + components +
    \texttt{offmanifold\_residual} to \texttt{metrics\_all.csv}.
  \item Validate end-to-end on the offline synthetic planted-concept smoke task; confirm
    \texttt{onmanifold\_steer} beats \texttt{naive}/\texttt{random} on synthetic-CFS.
\end{enumerate}

\begin{delivbox}[Deliverable \#2 --- Pipeline \& Implementation]
A runnable pipeline: frozen CLIP $\to$ TopK SAE $\to$ concept directions $\to$ the four
steering variants + supervised reference $\to$ CFS scoring, with all five OOD datasets
loadable and the synthetic smoke path green.
\end{delivbox}
\begin{checkbox}[Acceptance Check]
\texttt{python -m src.run\_experiments --smoke} writes \texttt{metrics\_all.csv} with a
row per variant including \texttt{cfs} and \texttt{offmanifold\_residual};
\texttt{onmanifold\_steer} shows lower off-manifold residual and higher synthetic-CFS
than \texttt{naive}/\texttt{random}; the real CLIP + ImageNet-shift loaders import and
return correctly-shaped batches on at least one shift level.
\end{checkbox}

% ===========================================================================
\section{Milestone 3: Training \& Ablation Experiments}

\renewcommand{\arraystretch}{1.2}
\begin{center}\small
\begin{tabular}{@{}p{0.30\linewidth} p{0.26\linewidth} p{0.36\linewidth}@{}}
\toprule
\textbf{Model / Steering method} & \textbf{Cache-or-cost} & \textbf{Primary metrics}\\
\midrule
Supervised TCAV-style (reference) & Label-expensive direction & CFS + components
  (the gold to approach)\\
\texttt{onmanifold\_steer} (ours) & $P_M\Delta$, rank $r$ & CFS, off-manifold residual,
  OOD slope\\
\texttt{clamp\_steer} & Fixed-magnitude clamp, no projection & CFS, residual\\
\texttt{naive\_steer} & Off-manifold $a+s\,d$ & CFS, residual (main competitor)\\
\texttt{random\_steer} & Random direction & CFS (null/sanity floor)\\
\bottomrule
\end{tabular}
\end{center}

\textbf{Implementation details.} All methods are evaluated at matched steering strength
on the same selected concept set; readout probes are held out from selection; the OOD
ladder uses identical preprocessing per shift; CFS, its components, and the off-manifold
residual are bootstrapped over concepts. The core experiment is the $r\times s$ design
grid (A3 $\times$ steering-strength) feeding the headline CFS-vs-severity curve.

\subsection*{Error taxonomy}
\renewcommand{\arraystretch}{1.2}
\begin{center}\small
\begin{tabular}{@{}p{0.24\linewidth} p{0.46\linewidth} p{0.24\linewidth}@{}}
\toprule
\textbf{Error Type} & \textbf{Description} & \textbf{Expected Frequency}\\
\midrule
Off-Manifold Artifact & Naive edit leaves the real-image manifold; readout moves but the
  change is non-realistic / undecodable (a fake causal effect) & High for
  \texttt{naive\_steer}/\texttt{clamp\_steer}; rare for \texttt{onmanifold\_steer}\\
Specificity Leakage & Steering the target concept also moves off-target probes
  (entangled / polysemantic direction) & Medium; rises with steering strength\\
Polysemantic / Uninterpretable Feature & The SAE feature encodes several concepts (or
  none clean) $\Rightarrow$ no clean knob; filtered by A4 & Medium pre-selection;
  $\sim$85\% of raw features\\
OOD Collapse & Faithful in-distribution but CFS collapses under shift
  (Sketch / C-high / ObjectNet) --- the headline failure mode & Rises sharply past the
  shift knee\\
Magnitude Saturation & Beyond a strength threshold the readout flattens / clips;
  monotonicity breaks & Rises with steering strength\\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Ablations A1--A5}
\begin{description}[leftmargin=2.6em,style=nextline,itemsep=2pt]
  \item[A1 --- SAE type] TopK vs.\ L1 (vanilla) SAE: effect on clean concept magnitudes
    and CFS.
  \item[A2 --- TopK $k$] Sparsity-level sweep: how $k$ trades interpretability against
    faithfulness; CFS per $k$.
  \item[A3 --- Projection rank $r$] The core knob: low $r$ over-constrains (effect dies),
    high $r$ lets the edit drift off-manifold; locate the CFS knee.
  \item[A4 --- Concept-selection threshold] How strict the ``well-defined concept''
    filter is, and the resulting reliable-concept fraction (the $\sim$10--15\% tail);
    CFS per threshold.
  \item[A5 --- Backbone layer \& token] Which ViT-B/16 layer, and patch vs.\ CLS tokens,
    for SAE training and steering; CFS per choice.
\end{description}
The core ablation is the \textbf{A3 $\times$ steering-strength} 2-D grid
(figures \texttt{fig5}/\texttt{fig3}).

\begin{delivbox}[Deliverable \#3 --- Experiments \& Ablations]
A complete \texttt{results/metrics\_all.csv} covering all 5 steering methods across the
full OOD ladder, the A1--A5 ablation grids (with the A3$\times s$ core grid), and a
populated error-taxonomy tally --- all with bootstrap CIs over concepts.
\end{delivbox}
\begin{checkbox}[Acceptance Check]
Every (method, shift level) cell has a CFS with a confidence interval;
\texttt{random\_steer} sits at the floor and the supervised reference near the top
(sanity); the A3 sweep shows a non-monotone CFS-vs-$r$ curve with an identifiable knee;
the strength sweep shows the saturation onset; each error type has a measured frequency.
\end{checkbox}

% ===========================================================================
\section{Milestone 4: Analysis, Insights \& Manuscript}

\textbf{Core visualizations (9).}
\begin{itemize}[leftmargin=1.4em,itemsep=0pt]
  \item \texttt{fig1\_cfs\_ood\_sweep} --- \textbf{headline}: CFS vs.\ shift severity
    (clean $\to$ R $\to$ Sketch $\to$ C-1..5 $\to$ ObjectNet), on-manifold vs.\ naive;
    collapse knee marked, CI band.
  \item \texttt{fig2\_faithfulness\_pareto} --- specificity vs.\ effect-size across all 5
    methods; on-manifold on the frontier, random in the corner.
  \item \texttt{fig3\_strength\_sweep} --- CFS vs.\ strength $s$ with the saturation knee.
  \item \texttt{fig4\_concept\_reliability} --- per-concept CFS distribution: the long
    unreliable mass and the $\sim$10--15\% reliable tail.
  \item \texttt{fig5\_ood\_heatmap} --- shift type $\times$ method CFS grid; best cell boxed.
  \item \texttt{fig6\_monotonicity\_curve} --- example readout vs.\ knob: smooth
    (on-manifold) vs.\ jagged (naive).
  \item \texttt{fig7\_by\_method\_bar} --- mean CFS by variant with bootstrap CIs; plus
    schematics \texttt{fig\_overview} and \texttt{fig\_method}.
\end{itemize}

\textbf{Insight report.} \emph{RQ1:} whether on-manifold beats naive / random / clamp at
matched strength on clean images and how close it gets to the supervised reference, with
CIs. \emph{RQ2:} decompose CFS into monotonicity / specificity / sufficiency, give the
reliable-concept fraction against the $\sim$10--15\% claim, and locate the optimum over
$(s,r)$. \emph{RQ3:} the CFS-vs-severity curve, the $\Delta$CFS-per-shift slope, the
collapse knee, and whether on-manifold degrades more gracefully than naive.

\renewcommand{\arraystretch}{1.2}
\begin{center}\small
\begin{tabular}{@{}p{0.27\linewidth} p{0.30\linewidth} p{0.36\linewidth}@{}}
\toprule
\textbf{Use Case} & \textbf{Recommended Design} & \textbf{Reasoning}\\
\midrule
Trustworthy concept editing on clean data & \texttt{onmanifold\_steer}, moderate $r$,
  pre-saturation $s$ & Highest clean CFS at matched strength; low off-manifold residual\\
Auditing under domain shift & On-manifold + report CFS at each shift level & Tracks the
  collapse knee; flags where a ``concept'' stops being causal\\
Label-rich, accuracy-critical concept & Supervised TCAV-style reference & The expensive
  gold; use when labels exist and faithfulness is paramount\\
Quick unsupervised exploration & On-manifold over a strict A4-filtered concept set & Skip
  the $\sim$85\% polysemantic tail; spend effort on reliable knobs\\
Never & \texttt{naive\_steer}/\texttt{random\_steer} as evidence & High apparent effect,
  low real faithfulness --- artifact or null\\
\bottomrule
\end{tabular}
\end{center}

\textbf{Manuscript outline (8 sections, IEEE conference).}
(1)~Introduction \& the open gap; (2)~Background (vision SAEs, steering, faithfulness,
robustness); (3)~Method --- the edit $a'=a+s(P_M\Delta)$ and the CFS definition;
(4)~Experimental setup; (5)~Results --- RQ1--RQ3 with the 9 figures; (6)~Ablations
A1--A5; (7)~Discussion, limitations \& the per-concept reliability story;
(8)~Conclusion. \textbf{Packaging:} reproducible config, \texttt{metrics\_all.csv},
figure scripts, and data cards, all consistent with this roadmap and the brief.

\begin{delivbox}[Deliverable \#4 --- Analysis \& Manuscript]
The 9 figures rendered from real M3 results; an insight report answering RQ1--RQ3; the
practitioner decision table; and a complete IEEE-conference manuscript draft.
\end{delivbox}
\begin{checkbox}[Acceptance Check]
Each RQ has a one-paragraph, evidence-backed answer with CIs; every figure caption
states the measured (not illustrative) numbers; the decision table's recommendations
follow from the results; the manuscript compiles and matches this roadmap and the brief
on title, RQs, baselines, benchmarks, and variant names.
\end{checkbox}

% ===========================================================================
\section{One-Page Checklist}

\begin{multicols}{2}
\textbf{Weeks 1--2 --- Literature \& Design}
\begin{itemize}[leftmargin=1.4em,label={},itemsep=0pt]
  \item \cbox Read \& annotate all 12 papers; draft the $\sim$24-entry bibliography.
  \item \cbox Fix the CFS definition and the on-manifold edit $a'=a+s(P_M\Delta)$.
  \item \cbox Specify the 4 variants + supervised reference; matched-strength protocol.
  \item \cbox Deliverable~\#1 + Acceptance Check signed off.
\end{itemize}

\textbf{Weeks 3--4 --- Data \& Implementation}
\begin{itemize}[leftmargin=1.4em,label={},itemsep=0pt]
  \item \cbox CLIP ViT-B/16 activations cached; TopK SAE trained; $U_r$ estimated.
  \item \cbox Four steering variants + readout / off-target probes implemented.
  \item \cbox OOD loaders (R / Sketch / C / ObjectNet) and CFS dispatcher wired.
  \item \cbox Synthetic smoke green; Deliverable~\#2 + Acceptance Check.
\end{itemize}

\columnbreak

\textbf{Weeks 5--6 --- Experiments \& Ablations}
\begin{itemize}[leftmargin=1.4em,label={},itemsep=0pt]
  \item \cbox CFS for all 5 methods across the full OOD ladder, bootstrapped.
  \item \cbox Ablations A1--A5, including the A3$\times s$ core grid.
  \item \cbox Error taxonomy tallied; \texttt{metrics\_all.csv} complete.
  \item \cbox Deliverable~\#3 + Acceptance Check.
\end{itemize}

\textbf{Weeks 7--8 --- Analysis \& Manuscript}
\begin{itemize}[leftmargin=1.4em,label={},itemsep=0pt]
  \item \cbox All 9 figures rendered from real results.
  \item \cbox Insight report answers RQ1--RQ3 with CIs; decision table filled.
  \item \cbox IEEE-conference manuscript drafted; consistent with the brief.
  \item \cbox Deliverable~\#4 + Acceptance Check; repository packaged.
\end{itemize}
\end{multicols}

% ===========================================================================
\section{Quick Reference Links}

\begin{multicols}{2}
\begin{itemize}[leftmargin=1.2em,itemsep=0pt]
  \item Cunningham et al., \emph{SAEs Find Interpretable Features} ---
    \href{https://arxiv.org/abs/2309.08600}{2309.08600}
  \item Gao et al., \emph{Scaling \& Evaluating SAEs (TopK)} ---
    \href{https://arxiv.org/abs/2406.04093}{2406.04093}
  \item Turner et al., \emph{Activation Addition} ---
    \href{https://arxiv.org/abs/2308.10248}{2308.10248}
  \item Zou et al., \emph{Representation Engineering} ---
    \href{https://arxiv.org/abs/2310.01405}{2310.01405}
  \item Kim et al., \emph{TCAV} ---
    \href{https://arxiv.org/abs/1711.11279}{1711.11279}
  \item Radford et al., \emph{CLIP} ---
    \href{https://arxiv.org/abs/2103.00020}{2103.00020}
  \item Dosovitskiy et al., \emph{ViT} ---
    \href{https://arxiv.org/abs/2010.11929}{2010.11929}
  \item Hendrycks et al., \emph{ImageNet-R} ---
    \href{https://arxiv.org/abs/2006.16241}{2006.16241}
  \item Hendrycks \& Dietterich, \emph{ImageNet-C} ---
    \href{https://arxiv.org/abs/1903.12261}{1903.12261}
  \item Wang et al., \emph{ImageNet-Sketch} ---
    \href{https://arxiv.org/abs/1905.13549}{1905.13549}
  \item Makelov et al., \emph{Principled SAE Eval} ---
    \href{https://arxiv.org/abs/2405.08366}{2405.08366}
  \item Marks et al., \emph{Sparse Feature Circuits} ---
    \href{https://arxiv.org/abs/2403.19647}{2403.19647}
\end{itemize}
\end{multicols}

\end{document}
